---
title: |
   | Visual Search Familiar Faces:
   | Figures and Descriptive Statistics for Accuracy
author: "Matteo Visconti di Oleggio Castello"
output: 
    pdf_document:
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

# Figures and Descriptive Statistics for Accuracy

This R Markdown file produces descriptive statistics and figures for accuracy.

## Setup

Load some useful libraries and load the data.
```{r}
require(dplyr)
require(ggplot2)
require(doParallel)
require(foreach)
require(knitr)

data <- read.csv('data.csv')
```

Set up some variables that will be used later.
```{r}
nproc <- 4  # change this to use more/less processors for parallel use
seed <- 42  # seed for rng to obtain reproducible results in different runs
alpha <- .05  # significance level for confidence intervals
nbs <- 10000  # number of bootstrapping repetitions
```

## Bootstrapping of the data

Bootstrap data resampling within each condition. 
It could take a while depending on the number of processors used.

```{r, results='hide', cache=TRUE}
cl <- makeCluster(nproc)
registerDoParallel(cl)
# push required packages to each worker
clusterCall(cl, function() library(magrittr))
clusterCall(cl, function() library(plyr))
clusterCall(cl, function() library(dplyr))
bstrap <- data.frame()
set.seed(seed)
bstrap <- foreach(i = 1:nbs, .combine=rbind) %dopar% {
    tmp <- 
        data %>%
        group_by(orientation, target_presence, familiarity, 
                 set_size, target_sex, subid) %>%
        sample_frac(1, replace=T) %>%
        group_by(orientation, target_presence, familiarity, set_size) %>%
        summarise(accuracy=sum(correct == 1)/n() * 100)
    
    tmp$index <- i
    tmp
}
stopCluster(cl)
```

Visualize bootstrapping distribution

```{r, fig.height=16, fig.width=12}
bstrap$orientation <- factor(bstrap$orientation, 
                             levels=c('Upright', 'Inverted'))
bstrap$target_presence <- factor(bstrap$target_presence, 
                              levels=c('Target Present', 'Target Absent'))

ggplot(bstrap, aes(accuracy, color=familiarity)) + 
    geom_freqpoly(binwidth=.5) + 
    facet_grid(target_presence*set_size ~ orientation) +
    theme_bw(base_size=12) +
    labs(color='Target', x='Accuracy [%]', y='Count') +
    scale_color_brewer(palette='Set1')
```

## Plot of Average Accuracy
 Now compute confidence intervals and averages from the bootstrapped samples.

```{r}
cis <- 
    bstrap %>% 
    group_by(orientation, target_presence, familiarity, set_size) %>%
    summarise(low=quantile(accuracy, alpha/2), 
              high=quantile(accuracy, 1-alpha/2))

# compute average from original data
avgs <- data %>%
    group_by(orientation, target_presence, familiarity, set_size) %>%
    summarise(accuracy=sum(correct == 1)/n() * 100)

cis <- merge(avgs, cis)
```

Now plot the figure

```{r}
# first reorder the levels of target_presence and orientation
cis$orientation <- factor(cis$orientation, levels=c('Upright', 'Inverted'))
cis$target_presence <- factor(cis$target_presence, 
                              levels=c('Target Present', 'Target Absent'))

pd = position_dodge(width=1)

ggplot(cis, aes(set_size, accuracy, ymin=low, ymax=high, 
                color=familiarity)) + 
    geom_line(alpha=.8, size=1, position=pd) +
    geom_errorbar(width=0.3,  position=pd)  + 
    geom_point(position=pd) +
    facet_grid(target_presence ~ orientation) +
    theme_bw(base_size=12) +
    theme(panel.grid.major.y = element_line(colour = "gray30"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.y = element_blank()) +
    scale_x_continuous(breaks=c(2, 4, 6)) +
    labs(x='Set Size', y='Accuracy [%]', color='Target', 
         linetype='Orientation') +
    scale_color_brewer(palette='Set1')
```

These are the plotted values
```{r}
kable(cis, digits=2)
```

## Compute Accuracy across conditions

Accuracy according to target presence. 
```{r}
accs_bytp <- 
    avgs %>%
    group_by(target_presence) %>%
    summarise(accuracy=mean(accuracy))

cis_bytp <- 
    bstrap %>% group_by(index, target_presence) %>%
    summarise(accuracy=mean(accuracy)) %>%
    group_by(target_presence) %>%
    summarise(low=quantile(accuracy, alpha/2), 
              high=quantile(accuracy, 1-alpha/2))
bytp <- merge(accs_bytp, cis_bytp)

kable(bytp, digits=2)
```

Accuracy according to target orientation.
```{r}
accs_byto <- 
    avgs %>%
    group_by(orientation) %>%
    summarise(accuracy=mean(accuracy))
cis_byto <- 
    bstrap %>% group_by(index, orientation) %>%
    summarise(accuracy=mean(accuracy)) %>%
    group_by(orientation) %>%
    summarise(low=quantile(accuracy, alpha/2), 
              high=quantile(accuracy, 1-alpha/2))

byto <- merge(accs_byto, cis_byto)

kable(byto, digits=2)
```
    
Accuracy according to set size.
```{r}
accs_byss <- 
    avgs %>%
    group_by(set_size) %>%
    summarise(accuracy=mean(accuracy))
cis_byss <- 
    bstrap %>% group_by(index, set_size) %>%
    summarise(accuracy=mean(accuracy)) %>%
    group_by(set_size) %>%
    summarise(low=quantile(accuracy, alpha/2), 
              high=quantile(accuracy, 1-alpha/2))

byss <- merge(accs_byss, cis_byss)

kable(byss, digits=2)
```

Accuracy according to familiarity.
```{r}
accs_byfam <- 
    avgs %>%
    group_by(familiarity) %>%
    summarise(accuracy=mean(accuracy))
cis_byfam <- 
    bstrap %>% group_by(index, familiarity) %>%
    summarise(accuracy=mean(accuracy)) %>%
    group_by(familiarity) %>%
    summarise(low=quantile(accuracy, alpha/2), 
              high=quantile(accuracy, 1-alpha/2))

byfam <- merge(accs_byfam, cis_byfam)
kable(byfam, digits=2)
```

Accuracy according to target presence and set size.
```{r}
accs_byssXtp <-
    avgs %>%
    group_by(target_presence, set_size) %>%
    summarise(accuracy=mean(accuracy))
cis_byssXtp <-
    bstrap %>% group_by(index, set_size, target_presence) %>%
    summarise(accuracy=mean(accuracy)) %>%
    group_by(target_presence, set_size) %>%
    summarise(low=quantile(accuracy, alpha/2), 
              high=quantile(accuracy, 1-alpha/2))

byssXtp <- merge(accs_byssXtp, cis_byssXtp)

kable(byssXtp, digits=2)
```

Accuracy according to target orientation and target presence.
```{r}
accs_bytoXtp <-
    avgs %>%
    group_by(target_presence, orientation) %>%
    summarise(accuracy=mean(accuracy))
cis_bytoXtp <-
    bstrap %>% group_by(index, target_presence, orientation) %>%
    summarise(accuracy=mean(accuracy)) %>%
    group_by(target_presence, orientation) %>%
    summarise(low=quantile(accuracy, alpha/2), 
              high=quantile(accuracy, 1-alpha/2))

bytoXtp <- merge(accs_bytoXtp, cis_bytoXtp)

kable(bytoXtp, digits=2)
```

Accuracy according to familiarity and target presence.

```{r}
accs_byfamXtp <-
    avgs %>%
    group_by(target_presence, familiarity) %>%
    summarise(accuracy=mean(accuracy))
cis_byfamXtp <-
    bstrap %>% group_by(index, target_presence, familiarity) %>%
    summarise(accuracy=mean(accuracy)) %>%
    group_by(target_presence, familiarity) %>%
    summarise(low=quantile(accuracy, alpha/2), 
              high=quantile(accuracy, 1-alpha/2))

byfamXtp <- merge(accs_byfamXtp, cis_byfamXtp)

kable(byfamXtp, digits=2)
```

