---
title: |
   | Familiarity Facilitates Feature-based Face Processing
   | Models for Accuracy
author: "Matteo Visconti di Oleggio Castello, Kelsey G. Wheeler, Carlo Cipolli, M. Ida Gobbini"
output: 
    pdf_document:
        toc: true
---

# Setup

In this document we'll analyze the accuracy creating Logit Mixed-Effect
Models separately for Target Present and Target Absent trials.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
# return version information
version
packages <- c('lme4',
              'car',
              'dplyr')

for (package in packages) {
  require(package, character.only=T)
  cat(paste(package, packageVersion(package), '\n'))
}

data <- read.csv('../data/data.csv')
# set order of levels for plotting
data$orientation <- factor(data$orientation, 
                           levels=c('Upright', 'Inverted'))
data$target_presence <- factor(data$target_presence, 
                               levels=c('Target Present', 'Target Absent'))

# set set_size as a factor
data$set_size <- as.factor(data$set_size)
```

Set up zero-sum contrasts for factors.

```{r}
contrasts(data$set_size) <- contr.poly(3)
contrasts(data$orientation) <- c(-1, 1)
contrasts(data$familiarity) <- c(-1,1)
contrasts(data$target_sex) <- c(-1,1)
```

# Model on Target Present Trials

```{r}
# get target present trials
data_tp <- data %>%
    filter(target_presence == 'Target Present')
```

Try to fit a very general model using the `bobyqa` optimizer (`Nelder_Mead` is
another option but might not converge sometimes). 
```{r}
m1 <- glmer(correct ~ set_size*familiarity*orientation + target_sex + 
               (1 + target_sex | subid) + 
               (1 + target_sex | stimuli_combination),
           family=binomial,
           data=data_tp,
           control=glmerControl(optimizer="bobyqa"))
```

Now we start reducing the random effect structure. First removing the random
slope for `stimuli_combination`.

```{r}
m2 <- glmer(correct ~ set_size*familiarity*orientation + target_sex + 
               (1 + target_sex | subid) + 
               (1 | stimuli_combination),
           family=binomial,
           data=data_tp,
           control=glmerControl(optimizer="bobyqa"))

anova(m1, m2)
```

The two models are not statistically different, thus we keep reducing `m2` by
entirely removing the random effect for `stimuli_combination`.

```{r}
m3 <- glmer(correct ~ set_size*familiarity*orientation + target_sex + 
               (1 + target_sex | subid),  
           family=binomial,
           data=data_tp,
           control=glmerControl(optimizer="bobyqa"))

anova(m2, m3)
```

The random effect for `stimuli_combination` doesn't seem to be necessary. We
thus try to reduce model `m3` further by removing the random slope for 
`subid`.

```{r}
m4 <- glmer(correct ~ set_size*familiarity*orientation + target_sex + 
               (1 | subid),  
           family=binomial,
           data=data_tp,
           control=glmerControl(optimizer="bobyqa"))

anova(m3, m4)
```

We thus keep model `m4` as our final model. Now we'll test significance of the 
factors using Type 3 Analysis of Deviance with Wald's $\chi^2$ test.
```{r}
Anova(m4, type=3)
```

# Model on Target Absent Trials
We'll repeat the same process for Target Absent trials.
```{r}
# get target absent trials
data_ta <- data %>%
    filter(target_presence == 'Target Absent')
```

Fit a very general model.
```{r}
m1 <- glmer(correct ~ set_size*familiarity*orientation + target_sex + 
               (1 + target_sex | subid) + 
               (1 + target_sex | stimuli_combination),
           family=binomial,
           data=data_ta,
           control=glmerControl(optimizer="bobyqa"))
```

Remove the random slope for `stimuli_combination`.

```{r}
m2 <- glmer(correct ~ set_size*familiarity*orientation + target_sex + 
               (1 + target_sex | subid) + 
               (1 | stimuli_combination),
           family=binomial,
           data=data_ta,
           control=glmerControl(optimizer="bobyqa"))

anova(m1, m2)
```

The two models are not statistically different, thus we keep reducing `m2` by
entirely removing the random effect for `stimuli_combination`.

```{r}
m3 <- glmer(correct ~ set_size*familiarity*orientation + target_sex + 
               (1 + target_sex | subid),  
           family=binomial,
           data=data_ta,
           control=glmerControl(optimizer="bobyqa"))

anova(m2, m3)
```

The random effect for `stimuli_combination` doesn't seem to be necessary. We
thus try to reduce model `m3` further by removing the random slope for 
`subid`.

```{r}
m4 <- glmer(correct ~ set_size*familiarity*orientation + target_sex + 
               (1 | subid),  
           family=binomial,
           data=data_ta,
           control=glmerControl(optimizer="bobyqa"))

anova(m3, m4)
```

For Target Absent trials the random slope for `subid` significantly improves
the fit of the model. Thus, we keep `m3` as our final model.
We'll test significance of the factors using Type 3 Analysis of Deviance 
with Wald's $\chi^2$ test.
```{r}
Anova(m3, type=3)
```

