---
title: |
   | Visual Search Familiar Faces:
   | Figures and Effect Sizes for Reaction Times
author: "Matteo Visconti di Oleggio Castello"
output: 
    pdf_document:
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

# Figures, Descriptive Statistics, and Effect Sizes for Reaction Times

This R markdown file will produce the figures for the manuscript.

Load some useful libraries and load the data.
```{r}
require(dplyr)
require(ggplot2)
require(doParallel)
require(foreach)
require(assertthat)
require(knitr)

data <- read.csv('data.csv')
data_correct <- data %>% filter(correct == 1)
```

Set up some variables that will be used later.
```{r}
nproc <- 4  # change this to use more/less processors for parallel use
seed <- 42  # seed for rng to obtain reproducible results in different runs
alpha <- .05  # significance level for confidence intervals
nbs <- 10000  # number of bootstrapping repetitions
```

## Bootstrapping of the data

Bootstrap data resampling within each condition. 
It could take a while depending on the number of processors used.

```{r, results='hide', cache=TRUE}
cl <- makeCluster(nproc)
registerDoParallel(cl)
# push required packages to each worker
clusterCall(cl, function() library(magrittr))
clusterCall(cl, function() library(plyr))
clusterCall(cl, function() library(dplyr))
bstrap <- data.frame()
set.seed(seed)
bstrap <- foreach(i = 1:nbs, .combine=rbind) %dopar% {
    tmp <- 
        data_correct %>%
        group_by(orientation, target_presence, familiarity, 
                 set_size, target_sex, subid) %>%
        sample_frac(1, replace=T) %>%
        group_by(orientation, target_presence, familiarity, set_size) %>%
        summarise(avg=mean(RT))
    
    tmp$index <- i
    tmp
}
stopCluster(cl)
```

Visualize bootstrapping distribution

```{r, fig.height=16, fig.width=12}
bstrap$orientation <- factor(bstrap$orientation, 
                             levels=c('Upright', 'Inverted'))
bstrap$target_presence <- factor(bstrap$target_presence, 
                              levels=c('Target Present', 'Target Absent'))

ggplot(bstrap, aes(avg, color=familiarity)) + 
    geom_freqpoly(binwidth=10) + 
    facet_grid(target_presence*set_size ~ orientation) +
    theme_bw(base_size=12) +
    labs(color='Target', x='Reaction Time [ms]', y='Count') +
    scale_color_brewer(palette='Set1')
```

## Plot of Average Reaction Times

Now compute confidence intervals and averages from the bootstrapped samples.

```{r}
cis <- 
    bstrap %>% 
    group_by(orientation, target_presence, familiarity, set_size) %>%
    summarise(low=quantile(avg, alpha/2), high=quantile(avg, 1-alpha/2))

# compute average from original data
avgs <- data_correct %>%
    group_by(orientation, target_presence, familiarity, set_size) %>%
    summarise(avg=mean(RT))

cis <- merge(cis, avgs)
```

Now plot the figure

```{r}
# first reorder the levels of target_presence and orientation
cis$orientation <- factor(cis$orientation, levels=c('Upright', 'Inverted'))
cis$target_presence <- factor(cis$target_presence, 
                              levels=c('Target Present', 'Target Absent'))

ggplot(cis, aes(set_size, avg, ymin=low, ymax=high, 
                color=familiarity, linetype=orientation)) + 
    geom_line(alpha=.8, size=1) +
    geom_errorbar(width=0.3, linetype='solid')  + 
    geom_point() +
    facet_grid(~ target_presence) +
    theme_bw(base_size=12) +
    theme(panel.grid.major.y = element_line(colour = "gray30"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.y = element_blank()) +
    scale_y_continuous(breaks=c(750, 1000, 1250, 1500, 1750)) +
    scale_x_continuous(breaks=c(2, 4, 6)) +
    labs(x='Set Size', y='RT [ms]', color='Target', linetype='Orientation') +
    scale_color_brewer(palette='Set1')
```

## Plot of Unstandardized Effect Sizes for each Set Size and Condition


Compute difference Stranger - Familiar for each bootstrapped sample and obtain
their confidence intervals.
```{r}
bstrap_fam <- filter(bstrap, familiarity == 'Familiar') %>% 
    select(-familiarity) 
bstrap_str <- filter(bstrap, familiarity == 'Stranger') %>% 
    select(-familiarity) 
# assert that the order is the same
tocheck <- c('orientation', 'target_presence', 'set_size', 'index')
for (check in tocheck) {
    assert_that(all(bstrap_fam[, check] == bstrap_str[, check]))
}

bstrap_es <- bstrap_str
bstrap_es$avg <- bstrap_es$avg - bstrap_fam$avg

# compute cis for effect size
cis_es <- 
    bstrap_es %>% 
    group_by(orientation, target_presence, familiarity, set_size) %>%
    summarise(low=quantile(avg, alpha/2), high=quantile(avg, 1-alpha/2))
```

Now we do the same in the original data.
```{r}
avg_fam <- avgs %>% filter(familiarity == 'Familiar') %>% select(-familiarity)
avg_str <- avgs %>% filter(familiarity == 'Stranger') %>% select(-familiarity)
# assert that the order is the same
tocheck <- c('orientation', 'target_presence', 'set_size')
for (check in tocheck) {
    assert_that(all(avg_fam[, check] == avg_str[, check]))
}

# get avg effect size
avg_es <- avg_str
avg_es$avg <- avg_es$avg - avg_fam$avg
```

Merge datasets and obtain data frame for plotting.
```{r}
cis_es <- merge(avg_es, cis_es)
cis_es$orientation <- factor(cis_es$orientation, 
                             levels=c('Upright', 'Inverted'))
cis_es$target_presence <- factor(cis_es$target_presence, 
                                 levels=c('Target Present', 'Target Absent'))
```


Plot the data.
```{r}
pd <- position_dodge(w=0.9)
ggplot(cis_es, aes(as.factor(set_size), avg, ymin=low, ymax=high, 
                   fill=orientation, group=orientation)) +
    geom_bar(position=pd, stat='identity', color='black') +
    geom_errorbar(width=0.2, position=pd) +
    facet_grid(~target_presence) +
    scale_fill_brewer(palette='Set2') +
    theme_bw() +
    labs(x='Set Size', y='Stranger - Familiar Effect Size [ms]', 
         fill='Orientation')
```

And these are the plotted data:
```{r}
kable(select(cis_es, -familiarity), digits=0)
```

## Unstandardized Effect Sizes across Set Size

Now we can compute the average effect size across set size.

```{r}
cis_es_setsize <- 
    bstrap_es %>%
    group_by(index, familiarity, orientation, target_presence) %>% 
    summarise(avg=mean(avg)) %>%
    group_by(orientation, target_presence) %>%
    summarise(low=quantile(avg, alpha/2), high=quantile(avg, 1-alpha/2))

avg_es_setsize <- 
    avg_es %>%
    group_by(orientation, target_presence) %>% 
    summarise(avg=mean(avg))

avg_es_setsize <- merge(avg_es_setsize, cis_es_setsize)

kable(avg_es_setsize, digits=0)
```

And also the average difference between Upright faces and Inverted faces.
```{r}
# for each bootstrap, obtain averages across set size
bstrap_es_setsize <- 
    bstrap_es %>%
    group_by(index, orientation, target_presence) %>%
    summarise(avg=mean(avg)) %>%
    ungroup() %>%
    arrange(index, target_presence, orientation) 

# subtract Inverted from Upright for every bootstrap and obtain CIs
cis_es_setsize_uprinv <- 
    bstrap_es_setsize %>%
    group_by(index, target_presence) %>%
    mutate(avg = avg - lag(avg)) %>%
    group_by(index, target_presence) %>%
    # remove rows with NA
    na.omit() %>%
    summarise(avg=avg) %>%
    group_by(target_presence) %>%
    summarise(low=quantile(avg, alpha/2), high=quantile(avg, 1-alpha/2))

# same for the original dataset
avg_es_setsize_uprinv <-
    avg_es_setsize %>%
    select(orientation, target_presence, avg) %>%
    ungroup() %>%
    arrange(target_presence, orientation) %>%
    group_by(target_presence) %>%
    mutate(avg = avg - lag(avg)) %>%
    select(-orientation) %>%
    na.omit()

avg_es_setsize_uprinv <-
    avg_es_setsize_uprinv %>% merge(cis_es_setsize_uprinv)

kable(avg_es_setsize_uprinv, digits=0)
```
