---
title: |
   | Familiarity Facilitates Feature-based Face Processing
   | Models for Reaction Times
author: "Matteo Visconti di Oleggio Castello, Kelsey G. Wheeler, Carlo Cipolli, M. Ida Gobbini"
output: 
    pdf_document:
        toc: true
---

# Setup

In this document we'll analyze the reaction times creating Linear Mixed-Effect
Models separately for Target Present and Target Absent trials.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
# return version information
version
packages <- c('lme4',
              'car',
              'dplyr')

for (package in packages) {
  require(package, character.only=T)
  cat(paste(package, packageVersion(package), '\n'))
}

data <- read.csv('data.csv')
# set order of levels for plotting
data$orientation <- factor(data$orientation, 
                           levels=c('Upright', 'Inverted'))
data$target_presence <- factor(data$target_presence, 
                               levels=c('Target Present', 'Target Absent'))
# get correct trials
data_correct <- data %>% filter(correct == 1)

# set set_size as a factor
data_correct$set_size <- as.factor(data_correct$set_size)
```

Set up zero-sum contrasts for factors.

```{r}
contrasts(data_correct$set_size) <- contr.poly(3)
contrasts(data_correct$orientation) <- c(-1, 1)
contrasts(data_correct$familiarity) <- c(-1,1)
contrasts(data_correct$target_sex) <- c(-1,1)
```

# Model on Target Present Trials

```{r}
# get target present trials
data_correct_tp <- data_correct %>%
    filter(target_presence == 'Target Present')
```

Try to fit a very general model: 
```{r}
m1 <- lmer(log(RT) ~ set_size*familiarity*orientation + target_sex + 
             (1 + target_sex | subid) + 
             (1 + target_sex | stimuli_combination),
           REML=F, data=data_correct_tp)
```
Model `m1` fails to converge, being very complex. We remove the random slope
for `target_sex` first from the random effects for `stimuli_combination`.

```{r}
m2 <- lmer(log(RT) ~ set_size*familiarity*orientation + target_sex + 
               (1 + target_sex | subid) + 
               (1 | stimuli_combination),
           REML=F, data=data_correct_tp)
```

Now reduce the complexity of the random effects structure and test it.
```{r}
m3 <- lmer(log(RT) ~ set_size*familiarity*orientation + target_sex + 
               (1 + target_sex | subid),
           REML=F, data=data_correct_tp)
anova(m3, m2)
```

Model `m2` is better, thus we'll keep the random effect for 
`stimuli_combination`. Let's remove the random slope for `subid` now.

```{r}
m4 <- lmer(log(RT) ~ set_size*familiarity*orientation + target_sex + 
               (1 | subid) + (1 | stimuli_combination),
           REML=F, data=data_correct_tp)
anova(m4, m2)
```

Model `m2` is still better, thus we'll keep its random effects structure.
We'll refit the model with Restricted Maximum Likelihood (REML), and check the
fitted vs. residual plot with and without log-transformation.

```{r}
m2_reml <- update(m2, REML=T)
m2_reml_nolog <- update(m2_reml, RT ~ .)

plot(m2_reml)
plot(m2_reml_nolog)
```

The log-transformation seems to work better than no transformation. Now we'll 
test significance of the factors using Type 3 Analysis of Deviance 
with Wald's $\chi^2$ test.
```{r}
Anova(m2_reml, type=3)
```

# Model on Target Absent Trials
```{r}
# get target absent trials
data_correct_ta <- data_correct %>%
    filter(target_presence == 'Target Absent')
```

Try to fit a very general model: 
```{r}
m1 <- lmer(log(RT) ~ set_size*familiarity*orientation + target_sex + 
             (1 + target_sex | subid) + 
             (1 + target_sex | stimuli_combination),
           REML=F, data=data_correct_ta)
```

Also in this case the general model `m1` fails. We'll thus follow the same 
steps as with target present trials. We remove the random slope
for `target_sex` first from the random effects for `stimuli_combination`.

```{r}
m2 <- lmer(log(RT) ~ set_size*familiarity*orientation + target_sex + 
               (1 + target_sex | subid) + 
               (1 | stimuli_combination),
           REML=F, data=data_correct_ta)
```

Now we reduce the complexity of the random effects structure and test it.
```{r}
m3 <- lmer(log(RT) ~ set_size*familiarity*orientation + target_sex + 
               (1 + target_sex | subid),
           REML=F, data=data_correct_ta)
anova(m3, m2)
```

Model `m2` is better, thus we'll keep the random effect for 
`stimuli_combination`. Let's remove the random slope for `subid` now.

```{r}
m4 <- lmer(log(RT) ~ set_size*familiarity*orientation + target_sex + 
               (1 | subid) + (1 | stimuli_combination),
           REML=F, data=data_correct_ta)
anova(m4, m2)
```

Model `m2` is still better, thus we'll keep its random effects structure. The
model thus has the same random effects structure as in the target present 
trials. We'll refit the model with Restricted Maximum Likelihood (REML), 
and test significance of the factors using Type 3 Analysis of Deviance 
with Wald's $\chi^2$ test.

```{r}
m2_reml <- update(m2, REML=T)

Anova(m2_reml, type=3)
```
